---
title: "Reading Data"
output: github_document
---

String vs. Factors
both look like character vectors, but
- strings are just strings
- factors have an underlying numeric structure with character labels sitting on top
factors generally make sense for variables that take on a few meaningful values

string
there are lots things can do with strings
- concatenating: joining snippets into a long
- shortening, subsetting, or turncating
- changing cases
- replacing one string segment with another 

factor
controlling factors is critical in several situations
- defining reference group in models
- ordering variables
- introducing new factor levels
common ooperation
- convert character variables to factors
- releveling by hand/count/a second variable
- renaming levels
- dropping unused levels 
the forcats package is the way to go for the majority of your factor needs

```{r setup}
library(tidyverse)
library(rvest)
```

## Strings and regex

```{r}
string_vec = c("my", "name", "is", "jeff")

str_detect(string_vec, "jeff")
str_detect(string_vec, "e") #does not have to match the pattern exactly
str_detect(string_vec, "Jeff") #uppercase and lowercase are different 

str_replace(string_vec,"jeff", "Jeff")
```

```{r}
string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )

str_detect(string_vec, "i think")
# can I detect strings that start with "i think"?
str_detect(string_vec, "^i think")
# ends with
str_detect(string_vec, "i think$")
```

```{r}
string_vec = c(
  "Y'all remember Pres. HW Bush?",
  "I saw a green bush",
  "BBQ and Bushwalking at Molonglo Gorge",
  "BUSH -- LIVE IN CONCERT!!"
  )

str_detect(string_vec, "bush")
str_detect(string_vec, "Bush")
# both "bush" and "Bush"
str_detect(string_vec, "[Bb]ush")
```

```{r}
string_vec = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )

#number first followed immediately by a letter 
str_detect(string_vec, "[0-9][a-zA-Z]")
str_detect(string_vec, "^[0-9][a-zA-Z]")
```

```{r}
string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )

str_detect(string_vec, "7.11")
#. -> special character matches anything
# to detect an actual dot:
str_detect(string_vec, "7\\.11")
```

```{r}
string_vec = c(
  'The CI is [2, 5]',
  ':-]',
  ':-[',
  'I found the answer on pages [6-7]'
  )

str_detect(string_vec, "\\[")
```

##factors

```{r}
factor_vec = factor(c("male","male","female", "female"))
# factors give levels (alphabet order)
# strings are in quotes

as.numeric(factor_vec)
```

relevel

```{r}
factor_vec = fct_relevel(factor_vec, "male")
# factor relevel: (a factor or character ) relevel by hand

as.numeric(factor_vec)
```

## NSDUH - strings

```{r}
url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

drug_use_html = read_html(url)

tabl_marj = 
  drug_use_html %>% 
  html_nodes(css = "table") %>% 
  first() %>%
  html_table() %>% #view()
  slice(-1) %>% #remove first row
  as_tibble()
```

```{r}
data_marj =
  tabl_marj %>% 
  select(-contains("P Value")) %>% 
  pivot_longer(
    -State, #eveything not in the state
    names_to = "age_year",
    values_to = "percent"
  ) %>% 
  separate(age_year, into = c("age","year"), sep = "\\(") %>% 
  mutate(
    year = str_replace(year, "\\)", ""), #replace ) with empty
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent)
  ) %>% 
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
```


## NSDUH - factor
```{r}
data_marj %>% 
  filter(age == "12-17") %>% 
  #mutate(State = fct_relevel(State, "Texas", "Oklahoma")) %>%
  mutate(State = fct_reorder(State, percent)) %>%  # the lowest percent will be the first; multiple percent observations per state will be calculated mean with the fct 
  ggplot(aes(State, y = percent, color = year)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

## Weather dataset

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

```{r}
weather_df %>% 
  mutate(name = fct_relevel(name, "Waikiki_HA")) %>% 
  #mutate(name = fct_order(name, tmax))
  ggplot(aes(x = name, y = tmax)) +
  geom_violin()
```

```{r}
weather_df %>% 
  lm(tmax ~ name, data= .)
```

